---
title: "Assignment 4 - Task 2"
author: "Jessica West"
date: "11/10/2021"
output: html_document
---

```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(ggbeeswarm)
library(broom)
library(effsize)
library(kableExtra)
library(devtools)
library(calecopal)
```

```{r}
bonanza_hares_juv_genders <- read_csv(here("data", "bonanza_hares.csv")) %>%
  filter(age == "j") %>% 
  mutate(sex = case_when(sex == "m" ~ "Male", 
                         sex == "f" ~ "Female",
                         sex %in% NA ~ "Unknown")) %>% 
  mutate(grid = case_when(grid == "bonbs" ~ "Bonaza Spruce", 
                         grid == "bonmat" ~ "Bonaza Mature",
                         grid == "bonrip" ~ "Bonaza Riparian"))
```

## Introduction

The "Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present" study explores the relationship between snowshoe hare (Lepus americanus) physical characteristic data (sex, age, and weight) of tagged individuals and population trends over the course of 14 years within the Bonanza Creek Experimental Forest, Alaska study area (Flora et al. 2002). Researchers trapped hares within the Tanana Valley in 3 study sites (Bonaza Spruce, Bonaza Mature, and Bonaza Riparian) to quantify hare density at a landscape scale. The study results showed no declines in survivorship despite a decline in density of the study populations. Additionally, body condition of individuals may be positively influencing the populations despite declining densities. This report analyzes the relationship between juvenile hare genders, body size, hind foot length.

## Data and Anaylses

Snowshoe hare structural size measurements were collected by Masters student B.K. Flora and team from the University of Alaska within the Bonanza Creek Experimental Forest study area (Flora et al. 2002). Data made available by Knut Kielland and colleagues, also from the University of Alaska Fairbanks (Kielland et al. 2002). The data contain observations for physical characteristic data (sex, age, and weight) collected between 1999 and 2012. Following exploratory data visualization, hind food lengths between male and female hares are by two-sample t-tests with a significance level ($\alpha$) of 0.05. Differences between variables are described by Cohen's *d* effect size. The relationship between hind foot length and body weight is analyzed via simple linear regression. All analyses are in R version 4.0.2 using RStudio version 1.3.1056. 

## Juvenile Hare Investigation

#### Annual Juvenile Hare Trap Counts

```{r}
hare_traps_year <- bonanza_hares_juv_genders %>% 
  mutate(date_new = mdy(date)) %>%
  mutate(year_new = year(date_new)) %>%
  group_by(year_new) %>%
  summarize(sample_size = n())

ggplot(data = hare_traps_year, aes(x = year_new, y = sample_size)) +
  geom_col(fill="skyblue3", color = "black", bins = 15) +
  scale_x_continuous(breaks = seq(from = 1999, to = 2012, by = 1)) +
  theme_gray() +
  labs(x = "Year",
       y = "Number of Hares",
       title = "Juvenile Hare Trapping Events",
       subtitle = "Bonanza Creek, Alaska: 1999-2012",
       caption = "Data source: Flora, B.K. (2002) & Kielland, K et al. (2017)")
```
```{r include = FALSE}

#Exploratory only

hare_traps_year %>% 
summarize(mean_sample_size = mean(sample_size, na.rm = TRUE),
          min_sample_size = min(sample_size, na.rm = TRUE),
          max_sample_size = max(sample_size, na.rm = TRUE))
summary(hare_traps_year)
```

**Figure 1:** The total number of juvenile snowshoe hares observed each year at the Bonanza Creek Long Term Ecological Research site between `r min(hare_traps_year$year_new)` and `r max(hare_traps_year$year_new)`. Data: Keilland et al. (2017).

The maximum number of juvenile hares trapped in one year during this period was `r max(hare_traps_year$n)` and the minimum number of juvenile hares trapped in one year was 0 during 2002 and 2009. The mean number of juvenile hares trapped per year is `r mean(hare_traps_year$n)`, and the median is `r median(hare_traps_year$n)`. A noticeable decline in juvenile trapping rates is visually apparent, indicating either a possible decline in population size (or possibly age class, though this cannot be determined without also looking at adult hare trap success), or the researchers simply had less overall success in trapping juvenile hares over the course of the study. Sampling efforts will need to be standardized in the future to acquire more individuals to provide a better understanding/ representation of the hare population. 

#### Juvenile Hare Weights

```{r}
juv_weight_genders <- bonanza_hares_juv_genders %>% 
  group_by(sex, grid)

ggplot(data = juv_weight_genders, aes(x = sex, y = weight, fill = sex)) +
  scale_fill_manual(values = cal_palette("sbchannel")) +
  geom_boxplot(outlier.color = "black") +
  geom_beeswarm(shape = 1, alpha = 0.75, aes(color = sex)) +
  scale_color_manual(values = c("black", "black", "black")) +
  facet_wrap(~grid) +
  labs(x = "Hare Gender",
       y = "Weight (g)",
       title = "Juvenile Hare Weights (Per Study Site)",
       subtitle = "Bonanza Creek, Alaska: 1999-2012",
       caption = "Data source: Flora, B.K. (2002) & Kielland, K et al. (2017)") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 2.5) +
  theme(legend.position = "none") +
  theme_gray()
```

**Figure 2:** Relationship between body weight (g) and gender of juvenile snowshoe hares observed in each study site within the Bonanza Creek Long Term Ecological Research study area between `r min(juv_weight_genders$year_new)` and `r max(juv_weight_genders$year_new)`. Individual data points are overlaid to show general trends, where box endpoints indicate the 25th and 75th percentile values. The middle black line and black diamond within the box indicate the average median and mean value for each gender. Whiskers correspond to the maximum (upper whisker) and minimum (lower whisker) values, excluding outliers (solid black dots). Data: Keilland et al. (2017)

Median and mean juvenile body weights (g) were higher for male snowshoe hares than female snowshoe hares at all three sites. Female hares averaged higher body weight values in the upper quartile (between the median and the 75th percentile), while male hares averaged higher body weight values in the lower quartile (between the median and the 25th percentile). Male and female body weights were highest at the Bonaza Spruce site, with higher mean and median weights than hares at the Bonaza Riparian or Bonaza Mature sites. Juvenile hares with unknown genders were only observed at the Bonanza Riparian and Bonanza Mature sites, with lower mean and median weights than male or female hares recorded at the Bonanza Riparian site.

#### Juvenile Weight Comparison (Males and Female)

```{r}
juv_weight_m_f_only <- bonanza_hares_juv_genders %>% 
  group_by(sex) %>% 
  filter(sex == "Male" | sex == "Female") %>% 
  summarize(mean_weight_g = mean(weight, na.rm = TRUE),
            sd_weight_g = sd(weight, na.rm = TRUE),
            sample_size = n())

juv_weight_m_f_only %>% 
    kable(digits = 2,
          caption = "Juvenile Hare Weights - Male and Female Only (1999-2012)", 
          col.names = c("Gender", "Mean Weight (g)", "Standard Deviation (g)", "Sample Size")) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```
**Table 1:** Relationship between body weight (g) and gender of juvenile male and female snowshoe hares observed in each study site within the Bonanza Creek Long Term Ecological Research study area between 1999 and 2012. One deceased hare (gender unknown) was recovered from the Bonanza Spruce site and no weight was taken. Data: Keilland et al. (2017). 

```{r include = FALSE}

#Exploratory graphs, tables, and stat tests only

ggplot(data = bonanza_hares_juv_genders, aes(x = weight)) +
   facet_wrap(~ sex) +
  geom_histogram(fill="lightskyblue2", color = "black", bins = 10) +
  theme_gray()

bonanza_hares_juv_genders %>%
  filter(sex %in% c("Male", "Female")) %>%
  group_by(sex) %>%
  summarize(Mean = mean(weight, na.rm = TRUE),
            Median = median(weight, na.rm  =TRUE), 
            Standard_Deviation = sd(weight, na.rm = TRUE), 
            Count = n()) %>% 
  kable(digits = c(2, 2, 2, 2), 
        col.names = c("Sex", "Mean (g)", "Median (g)", "Standard Deviation (g)", "Sample Size"), 
        align = "llcccc") %>%
  kable_styling(full_width = FALSE, position = "center", bootstrap_options = c("striped", "hover"))

ggplot(data = bonanza_hares_juv_genders, aes(x = weight)) +
  geom_histogram() + 
  facet_wrap(~ sex)

ggplot(data = bonanza_hares_juv_genders, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~ sex)

juv_hares_male <- bonanza_hares_juv_genders %>% 
  filter(sex == "Male") %>% 
  pull(weight) %>% 
  na.omit() 

juv_hares_female <- bonanza_hares_juv_genders %>% 
  filter(sex == "Female") %>% 
  pull(weight) %>% 
  na.omit() #

juv_hares_all <- c(juv_hares_female, juv_hares_male)

cohen_d_test <- cohen.d(juv_hares_male, juv_hares_female)

t_test <- t.test(juv_hares_male, juv_hares_female) 
```

Body weight was compared between hare genders where sampling site was not considered in this comparison (but may need to be in the future for a more accurate assessment of weights between sites). A Cohen's d test (`cohen.d()`) and a Welch two-sample t-test (`t.test()`) were run between male and female hare weights to determine the size and significance of gender on juvenile hare body weight. Diagnostic plots (not included) revealed the data to be normally distributed.

Mean weight for male hares (`r round(mean(juv_hares_male), 2)` g) is `r round(mean(juv_hares_male) - mean(juv_hares_female), 2)` g (`r round(abs(1-(mean(juv_hares_male)/mean(juv_hares_female)))*100, 2)`%) larger than mean weight for female hares (`r round(mean(juv_hares_female), 2)` g). Per the `cohen.d()` and `t.test()` results, gender and body weight are not closely related (Cohen's d = `r round(cohen_d_test$estimate, 2)`) though they are statistically significant (`r t_test$method`: p = `r round(t_test$p.value, 2)`). A p value of `r round(t_test$p.value, 2)` suggests that there is a `r round((t_test$p.value * 100), 2)` percent chance that if our male and female hare populations did have the same mean, the sample means would be at least `r round(mean(juv_hares_male) - mean(juv_hares_female), 2)` g different.

#### Relationship Between Juvenile Weight & Hind Foot Length

```{r}
ggplot(data = bonanza_hares_juv_genders, aes(x = hindft, y = weight)) +
  geom_point(color = "skyblue3") +
  theme_light() +
  labs(x = "Hind Foot Length (mm)",
       y = "Weight (g)",
       title = "Juvenile Hare Hind Foot Length and Weight",
       subtitle = "Bonanza Creek, Alaska: 1999-2012",
       caption = "Data source: Flora, B.K. (2002) & Kielland, K et al. (2017)")
```

**Figure 3:** Linear Relationship between hind foot length (mm) and body weight (g) for juvenile snowshoe hares observed within the Bonanza Creek Long Term Ecological Research study area between 1999 and 2012. Points indicate individual hare measurements. Data: Keilland et al. (2017).

```{r include = FALSE}

# Data analysis only

bonanza_hares_lm <- lm(weight ~ hindft, data = bonanza_hares_juv_genders)
hares_lm_tidy <- tidy(bonanza_hares_lm)
hares_lm_glance <- glance(bonanza_hares_lm)

hares_cor <- cor.test(bonanza_hares_juv_genders$weight, bonanza_hares_juv_genders$hindft)

hares_cor_tidy <- tidy(hares_cor)

plot(bonanza_hares_lm)

summary(bonanza_hares_lm)
```

```{r}
ggplot(data = bonanza_hares_juv_genders, aes(x = hindft, y = weight)) +
  geom_point(color = "skyblue3") +
  geom_smooth(method = "lm", color = "black") +
    theme_light() +
    labs(x = "Hind Foot Length (mm)",
       y = "Weight (g)",
       title = "Juvenile Hare Hind Foot Length and Weight",
       subtitle = "Bonanza Creek, Alaska: 1999-2012",
       caption = "Data source: Flora, B.K. (2002) & Kielland, K et al. (2017)")
```

**Figure 4:** Linear Relationship between hind foot length (mm) and body weight (g) for juvenile snowshoe hares observed within the Bonanza Creek Long Term Ecological Research study area between 1999 and 2012. Points indicate individual hare measurements. Linear model summary: $\beta$~1~ = `r round(hares_lm_tidy$estimate[2], 2)` g mm^-1^, p < 0.001, R^2^ = `r round(hares_lm_glance$r.squared,2)`, Pearson's *r* = `r round(hares_cor_tidy$estimate,2)`). Data: Keilland et al. (2017).

The relationship between body weight (g) and hind foot length (mm) appears relatively linear (Figure 4). Simple linear regression revealed that hind foot length significantly predicts body weight (p < 0.001, R^2^ = `r round(hares_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(hares_lm_tidy$estimate[2], 2)` g mm^-1^ (i.e., for each one millimeter increase in hind foot length we expect an average increase in body weight of `r round(hares_lm_tidy$estimate[2], 2)` g). Hind foot length and body weight are strongly, and significantly, positively correlated (Pearson's *r* = `r round(hares_lm_tidy$estimate,2)`, p < 0.001). Diagnostic plots (not included) reveal normally distributed and homoscedastic residuals. However, it is concerning that we are not incorporating gender (male or female) as possible influences of hind foot length and weight.

## Summary

Exploratory data analysis reveals the following initial findings: 

- Juvenile body weights (g) were higher for male snowshoe hares than female snowshoe hares at all three sites
- Male and female body weights were highest at the Bonaza Spruce site
- Hind foot length (mm) appears linearly related with body weight; the relationship (by simple linear regression) is significant, with a slope (expected average increase in body weight with a 1 mm increase in hind foot length) of `r round(hares_lm_tidy$estimate[2], 2)` g mm^-1^.

## Citations

Flora, B.K. 2002. Comparison of snowshoe hare populations in Interior Alaska. M.S. Thesis. University of Alaska-Fairbanks.

Kielland, Knut; Chapin, F Stuart; Ruess, Roger. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present, Bonanza Creek LTER - University of Alaska Fairbanks. BNZ:55, http://www.lter.uaf.edu/data/data-detail/id/55. doi:10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14
